---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r setoptions, results = "hide"}
library(knitr)
library(ggplot2)
library(dplyr)
library(mice)
opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data
```{r loadAndClean}
#download.file('https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip', destfile = 'activity.csv')

activity<- read.csv('activity.csv', header = TRUE)
activity <- activity[!is.na(activity$steps),]
activity$date <- as.Date(activity$date)
activity$steps <- as.numeric(as.character(activity$steps))
```


## What is mean total number of steps taken per day?
```{r plotDailySteps}
date_steps <- activity %>% group_by(date) %>% summarise(total_steps = sum(steps))

ggplot(date_steps) + geom_histogram(aes(x = date, weight = total_steps), binwidth = 1) + xlab("Date") + ylab("Steps") + ggtitle("Total Steps per Day")
```

## What is mean total number of steps taken per day?
```{r statsByDate}
date_stats <- activity %>% group_by(date) %>% summarise(total_steps = sum(steps), mean = mean(steps), median = median(steps))
print(date_stats, n = 53)
```

## What is the average daily activity pattern?
```{r plotTimeSeries}
ggplot(date_stats, aes(x = date, y = total_steps)) + geom_line() + xlab("Date") + ylab("Steps") + ggtitle("Time Series of Average Steps Taken")
```

#Time Interval with Maximum Number of Steps (on average)
```{r maxAverageStepsInterval}
interval_stats <- activity %>% group_by(interval) %>% summarise(total_steps = sum(steps), mean = mean(steps), median = median(steps))

max_avg_steps_in_interval <- max(interval_stats$total_steps)

max_interval <- interval_stats[interval_stats$total_steps == max_avg_steps_in_interval,]
```
##The interval with the maximum number of average steps is `r max_interval$interval`.

===============================================================

## Imputing missing values
```{r readDataAndLookForMissing}
activity<- read.csv('activity.csv', header = TRUE)
activity$steps <- as.numeric(as.character(activity$steps))
md.pattern(activity)
```

##Use **mice** package to impute data
```{r impute, cache = TRUE, results = "hide"}
impute_activity <- mice(data = activity, me = 5, method = "pmm", maxit = 50, seed = 500)

activity_imputed <- complete(impute_activity, 1)
```

#Total Number of Steps Taken Each Day After Imputing
```{r plotDailyStepsAfterImpute}
activity_imputed$date <- as.Date(activity_imputed$date)

ggplot(activity_imputed) + geom_histogram(aes(x = date, weight = steps), binwidth = 1) + xlab("Date") + ylab("Steps") + ggtitle("Total Steps per Day")
```

## Are there differences in activity patterns between weekdays and weekends?
```{r weekend}
activity_wkdy <- activity_imputed %>% mutate(wkday = weekdays(activity_imputed$date))

activity_wkdy <- activity_wkdy %>% mutate(wkday_type = ifelse(wkday %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"), "weekday", "weekend"))

interval_stats <- activity_wkdy %>% group_by(interval, wkday_type) %>% summarise(total_steps = sum(steps), mean = mean(steps), median = median(steps))

ggplot(interval_stats, aes(interval, total_steps)) + geom_line() + facet_grid(.~wkday_type) + ggtitle("Average Steps by Interval (Weekend vs. Weekdays)") + xlab("Interval") + ylab("Steps")
```